The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-01-11 MS_ToSubmit/RevisedSupMaterial.Rmd
Revised Sup Mat

The following figure and tables summarise the data we used to build the network of trophic (predator-prey) interactions for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem.
Figure 1 resumes the type and number of references reviewed. Table 1 is the complete list of trophic interactions and references that confirm each of them. Table 2 is the species list with details on trophic species cases (aggregated taxa).

```{r ref_plot, echo=FALSE, message=FALSE, fig.cap="Summary of references used to build the network of trophic (predator-prey) interactions for the MPAN-BB ecosystem."}

# Load packages
library(tidyverse)
library(ggplot2)

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, PreyGroup, PredGroup, Reference) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Figure 1. Plot resuming references
biblio <- data.frame(unique(int_good_res$Reference))
colnames(biblio) <- "Reference"
biblio <- biblio %>% 
  mutate(Type = case_when(str_detect(Reference, "pers.com.") ~ "PersComm",
                          str_detect(Reference, "etal") ~ "Article",
                          str_detect(Reference, "&") ~ "Article",
                          str_detect(Reference, "Encyclopedia of Life") ~ "DataBase",
                          str_detect(Reference, "GloBI") ~ "DataBase",
                          str_detect(Reference, "GATEWAy") ~ "DataBase",
                          str_detect(Reference, "PhDTesis") ~ "Thesis",
                          TRUE ~ "Article"))
ggplot(biblio) +
  geom_bar(aes(x = Type, fill = Type)) +
  theme_bw() +
  labs(x = "Reference", y = "Count") +
  annotate("text", x = Inf, y = Inf, label = paste("Total = ", nrow(biblio), sep = ""), vjust=1, hjust=1)

@@

```{r tableS1, echo=FALSE, message=FALSE, results='asis'}

@@

# Table 1. List of interactions
# landscape(knitr::kable(int_good_res %>% arrange(., (Predator), booktabs = TRUE, format = "latex", longtable = T, col.names = c("Prey","Predator","Reference", "Link"), caption = "List of predator-prey (trophic) interactions for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Alphabetically ordered by predator.") %>%
#              column_spec(1:3, width="2cm") %>% 
#              column_spec(4, width="14cm") %>% 
#              kable_styling(latex_options = c("hold_position","repeat_header"), font_size = 4))

pandoc.table(int_good_res %>% arrange(., (Predator)) %>% mutate(Prey = paste('\\tiny', Prey), Predator = paste('\\tiny', Predator), Reference = paste('\\tiny', Reference), Reference_link = paste('\\tiny', Reference_link)) %>% rename("Link"="Reference_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","right"),
             caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).")

@@

pandoc.table(sp_list %>% arrange(., (Species)) %>% mutate(Species = paste('\\tiny', Species), TaxLevel = paste('\\tiny', TaxLevel), FunctionalGroup = paste('\\tiny', FunctionalGroup), Reference = paste('\\tiny', Reference), Reference_pres_link = paste('\\tiny', Reference_pres_link), Observation = paste('\\tiny', Observation)) %>% rename("Taxonomy"="TaxLevel", "Group"="FunctionalGroup", "Link"="Reference_pres_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","right","right","left"),
             caption = "List of species for the food web of the Marine Protected Areas Namuncurá - Banco Burdwood I and II ecosystem. Alphabetically ordered by species.")

@@

```{r tableS3, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

@@

# Table 3. List of species
tbl_3 <- spp_total %>% 
  mutate(TopRole = case_when(TopRole == "hubcon" ~ "netcon", TRUE ~ TopRole)) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, TotalDegree, NumPrey, NumPred, Between, Close, TL, meanTrophicSimil, TopRole) %>% 
  arrange(., desc(TotalDegree)) %>% 
  mutate(across(where(is.numeric), round, 4),
         FunctionalGroup = case_when(TrophicSpecies == "Brachiopoda_spp" ~ "Brachiopoda",
                                     TrophicSpecies == "Campylaspis_spp" ~ "Cumacea",
                                     TRUE ~ FunctionalGroup))
# tbl_31 <- kableExtra::kbl(tbl_3, format = "latex", booktabs = TRUE, linesep = "", longtable = TRUE, col.names = c("Species","Group","Degree","NumPrey","NumPred","Between","Close","TL","TS","TR"), align = "llcccccccc", caption = "Species-level properties for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Ordered by degree (total number of interactions). Between: betweenness; Close: closeness; TL: trophic level; TS: trophic similarity; TR: topological role (netcon = network connector, modcon = module connector, modhub = module hub, modspe = module specialist).")
# tbl_32 <- kableExtra::kable_styling(tbl_31, latex_options = c("hold_position", "repeat_header"), font_size = 4.5)

@@

>>>>

The following figure and tables summarise the data we used to build the network of trophic (predator-prey) interactions for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem.
Figure 2 resumes the type and number of references reviewed. Table 1 is the complete list of trophic interactions and references that confirm each of them. Table 2 is the species list with details on trophic species cases (aggregated taxa).

With the gathered trophic data, we constructed a matrix of pairwise interactions; a value of 1 or 0 was assigned to each element a_ij of the matrix depending on whether the predator j preyed or not on the prey i. Then we transformed such a matrix into an oriented graph with L trophic interactions between S nodes or species (Figure 1).

```{r, out.width = "12cm", echo = FALSE, message=FALSE, fig.align = 'center', fig.cap = "Matrix of pairwise interactions and the oriented graph (food web) derived from it. Number of species S = 9, number of interactions (L) = 15."}
library(knitr)
knitr::include_graphics("FigS1.png")
```

### Network analysis

Trophic level

Trophic levels (TL) of species were calculated using the shortweighted TL formula of @Williams2004. Shortweighted trophic level is defined as the average of the shortest TL and prey-averaged TL. Shortest TL of a consumer in a food web is equal to 1 + the shortest chain length from this consumer to any basal species [@Williams2004]. Prey averaged TL is equal to 1 + the mean TL of all consumer's trophic resources, calculated as:

\begin{equation}
TL_j = 1 + \sum_{i = 1}^{S} l_{ij} \frac{TL_i}{n_j}
\end{equation}

where $TL_j$ is the trophic level of species $j$; $S$ is the number of species in the food web; $l_{ij}$ is the connection matrix with $S$ rows and $S$ columns, in which for column $j$ and row $i$, $l_{ij}$ = 1 if species $j$ consumes species $i$ and $l_{ij}$ = 0 if not; and $n_j$ is the number of prey species in the diet of species $j$. For basal species TL = 1 given that they have no prey.

Topological roles

Beforehand, we need to calculate modularity. Modularity was calculated as the difference between realized and expected within-modules interactions, divided by the total number of interactions. We used a stochastic algorithm called “simulated annealing” [@Guimera2005], that assumes that the nodes of the same module have more links than one would expect in a random network. The modules are obtained from dividing all the nodes in the network to maximize modularity. In this way, the modularity index is defined as:

\begin{equation}
M = \sum_{s} (\frac{I_s}{L} - (\frac{d_s}{2L})^2)
\end{equation}

where $s$ is the number of modules or compartments, $I_s$ is the number of links between species in the modules, $d_s$ is the sum of degrees for all species in module $s$ and $L$ is the total number of links.

Species topological roles were calculated using the method of functional cartography [@Guimera2005]. Roles are determined according to two parameters:

a) the standardized within-module degree (dz), a z-score that reflects how well a species is connected to other species inside the module, relative to other species within its own module:

\begin{equation}
dz_i = \frac{k_{is}-k_s} {ks}
\end{equation}

where $k_{is}$ is the number of links of species $i$ within its own module $s$, and $ks$ are the average and standard deviation of $k_{is}$ over all species in $s$, respectively.

b) the participation coefficient $PC$, estimates the links distribution of species $i$ among modules:

\begin{equation}
PC_i = 1 - \sum_{S} \frac{k_{is}}{k_i}
\end{equation}

where $k_i$ is the total number of links of species $i$ and $k_{is}$ is the number of links of species $i$ to species in module $s$.
To determine the role of each species, the $dz$ − $PC$ parameters space was divided into four regions [@Kortsch2015], with two threshold values: $dz$ = 2.5 and $PC$ = 0.625. Thus, the species were classified as follows:

* Module hub ($dz$ ≥ 2.5, $PC$ < 0.625): The species has a relatively high number of links, but at least 60% within its own module.

* Module specialist ($dz$ < 2.5, $PC$ < 0.625): The species has relatively few links and most within its own module.

* Module connector ($dz$ < 2.5, $PC$ ≥ 0.625): The species has relatively few links and most between modules.

* Network connector ($dz$ ≥ 2.5, $PC$ ≥ 0.625): The species has a high connectivity between and within the modules.

@@

```{r tableS1, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

@@

# Table S1. List of interactions
# knitr::kable(int_good_res %>% arrange(., (Predator)), booktabs = TRUE, format = "latex", col.names = c("Prey","Predator","Reference", "Link"), caption = "List of predator-prey (trophic) interactions for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Alphabetically ordered by predator.") %>%
#    column_spec(1:3, width="2cm") %>%
#    column_spec(4, width="14cm") %>%
#    kableExtra::kable_styling(latex_options = c("hold_position","repeat_header"), font_size = 4) %>%
#    kableExtra::landscape()

pandoc.table(int_good_res %>% arrange(., (Predator)) %>% mutate(Reference_link = paste('\\tiny', Reference_link), Reference = paste('\\tiny', Reference), Prey = paste('\\tiny', Prey), Predator = paste('\\tiny', Predator)) %>% rename("Link"="Reference_link"), style="multiline", split.tables = Inf, justify = c("left","left","center","right"),
             caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).")

@@

pandoc.table(sp_list %>% arrange(., (Species)) %>% mutate(Species = paste('\\tiny', Species), TaxLevel = paste('\\tiny', TaxLevel), FunctionalGroup = paste('\\tiny', FunctionalGroup), Reference = paste('\\tiny', Reference), Reference_pres_link = paste('\\tiny', Reference_pres_link), Observation = paste('\\tiny', Observation)) %>% rename("Taxonomy"="TaxLevel", "Group"="FunctionalGroup", "Link"="Reference_pres_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","left","left","left"),
             caption = "List of species for the food web of the Marine Protected Areas Namuncurá - Banco Burdwood I and II ecosystem. Alphabetically ordered by species.")

@@

```{r tableS3, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
require(kableExtra)

@@

# Table 3. List of species
tbl_3 <- spp_total %>% 
  mutate(TopRole = case_when(TopRole == "hubcon" ~ "netcon", TRUE ~ TopRole)) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, TotalDegree, NumPrey, NumPred, Between, Close, TL, meanTrophicSimil, TopRole) %>% 
  arrange(., desc(TotalDegree)) %>% 
  mutate(across(where(is.numeric), round, 4),
         FunctionalGroup = case_when(TrophicSpecies == "Brachiopoda_spp" ~ "Brachiopoda",
                                     TrophicSpecies == "Campylaspis_spp" ~ "Cumacea",
                                     TRUE ~ FunctionalGroup))

@@

## References
